---
title: "Homework 27"
format: html
editor: visual
---

Read through the code in this assignment. You will begin contributing to the code on line 36. Your tasks will be to:

-   Create a dataframe with just the complete cases

-   Review the 1-way ANOVA we ran during class with the full data set

-   Conduct a 2nd 1-way ANOVA using just the data set with complete cases and compare the results to the first analysis.

-   Conduct a linear regression to assess the relationship b/w Growth_rate_mm and Dried_mass_g, first with the complete data set and second with the complete cases data set and compare the results of the 2 analyses.

## Set up our workspace and read in the data

```{r}
rm(list = ls())
library(tidyverse)
library(ggfortify)
library(here)
```

Read in the data.

```{r}
df <- read.csv(here("Data/learn-about-NA.csv"))
str(df)
```

## Dealing with NAs in data analysis

Deciding how to treat NA values in your data can have critical impacts on your data analysis.

Let's look at an example.

Start by creating a data frame from df called df_All_NA, which includes the entire dataset including all NA values.

```{r}
df_All_NA <- df
```

Now use one of the methods we discussed in class on 13 November to create a second dataframe called `df_No_NA` that is just the complete cases of df.

```{r}
df_No_NA <- df[complete.cases(df), ] 
```

What if we wanted to test the hypothesis that there is no effect of fertilizer treatment on plant growth rate?

In class on Wednesday 13 November we ran a 1-way ANOVA on the complete data set. Let's review it here:

### First ANOVA using all the data

#### Plot and Guess

```{r}
ggplot(df_All_NA, aes(Treatment, Growth_rate_mm))+
  geom_boxplot()+
  geom_point()+
  theme_bw()
```

We decided that there was likely an effect of treatment on growth rate, but we weren't sure if both levels of the fertilizer treatment would differ from the control.

#### Create a model:

```{r}
grow.mod<-lm(Growth_rate_mm ~ Treatment, data = df_All_NA)
```

#### Check our assumptions:

```{r}
autoplot(grow.mod)
```

Looks pretty good.

#### Interpret results:

```{r}
anova(grow.mod)
```

Practice your interpretation.

Why are there 2 DF for the Treatment effect?

Answer: Because there are 3 levels of our Treatment variable: Control, Level 1, and Level 2. Number of groups - 1 = degrees of freedom therefore 3 - 1 = 2.

Why are there 46 DF for the Residuals? Remember that we determined that there were 49 values in the `Growth_rate_mm` variable using `sum(!is.na(variable))` like this: There are 46 DF for Residuals because you need to subtract the 3 treatment levels from 49.

```{r}
sum(!is.na(df$Growth_rate_mm))
```

So **there are only 49 data points used in the analysis**, even though the dataframe is 60 rows long!!!

49 - 3 = 46 = degrees of freedom for the residuals.

```{r}
summary(grow.mod)
```

Practice your interpretation:

Does adding 1 dose of fertilizer change the growth rate significantly from the Control? Yes

Does adding 2 doses of fertilizer change the growth rate significantly from the Control? Yes

We won't worry about re-plotting our data right now.

### Second Attempt: Complete Cases

Your turn. Run the 1-way ANOVA again, this time with only the dataset that has complete cases and that we called `df_No_NA`

#### Plot and Guess

```{r}
ggplot(df_No_NA, aes(Treatment, Growth_rate_mm))+
  geom_boxplot()+
  geom_point()+
  theme_bw()
```

What do you think? I think that there is likely an effect of treatment on growth rate, but we I'm not sure if both levels of the fertilizer treatment would be differet from the control.

#### Create a model:

```{r}
grow.mod.no.NA <- lm(Growth_rate_mm ~ Treatment, data = df_No_NA)
```

#### Check our assumptions:

```{r}
autoplot(grow.mod.no.NA)
```

What do you think? Same or different from our first 1-way ANOVA? I think that it is very similar to the first 1-way ANOVA we ran using NA values. There is a slight difference, mostly in the Residuals vs Leverage plot.

#### Interpret results:

```{r}
anova(grow.mod.no.NA)
```

Practice your interpretation.

Why are there 2 DF for the Treatment effect? There are 3 levels of our Treatment variable: Control, Level 1, and Level 2. Number of groups - 1 = degrees of freedom therefore 3 - 1 = 2.

Why are there 33 DF for the Residuals? You have to take the total number of observations (36) and subtract it by the number of variables being estimated.

How would you write a results statement? There is not a significant difference between Growth rate and Treatment level (F-value = 2.7906, df = 2).

```{r}
summary(grow.mod.no.NA)
```

Practice your interpretation:

Does adding 1 dose of fertilizer change the growth rate significantly from the Control? 1 dose of fertilizer does not significantly change the growth rate from the control.

Does adding 2 doses of fertilizer change the growth rate signficantly from the Control? 2 doses of fertilizer slightly significantly changes the growth rate from the control.

We won't worry about re-plotting our data right now.

## Linear Regression

Now conduct a linear regression to examine the relationship between `Growth_rate_mm` and `Dried_mass_g`, first with the complete data set and second with the complete cases data set and compare the results of the 2 analyses.

### First attempt: Full data set including NAs (use `df_All_NA`)

You know what to do here. Go for it.

#### Plot and guess

```{r}
ggplot(df_All_NA, aes(Growth_rate_mm, Dried_mass_g)) +
  geom_point() +
  theme_bw()
```

#### Create model

```{r}
plant.mod <- lm(Growth_rate_mm ~ Dried_mass_g, data = df_All_NA)
```

#### Check assumptions

```{r}
autoplot(plant.mod)
```

Don't worry if assumptions don't look perfect - focus on normality of the data.

#### Interpret model

Use `anova()` first before using `summary()`

```{r}
anova(plant.mod)
summary(plant.mod)
```

Why is there 1 degree of freedom for Growth_rate_mm and why are there 34 degrees of freedom for the Residuals?

There is 1 degree of freedom for Growth_rate_mm because there are two groups of variables, so you have to do the total number of groups subtracted by 1. So you get 2-1=1. There are 34 degrees of freedom for the Residuals because 24 observations out of the 60 were deleted due to missingness, as we see above, which leaves us with 36. Then you have to take the total number of observations (36) and subtract it by the number of variables being estimated. Which gives you 34.

### Second attempt: Limited data set with only complete cases (use `df_No_NA`)

#### Plot and guess

```{r}
ggplot(df_No_NA, aes(Growth_rate_mm, Dried_mass_g)) +
  geom_point() +
  theme_bw()
```

#### Create model

```{r}
plant.mod.no.na <- lm(Growth_rate_mm ~ Dried_mass_g, data = df_No_NA)
```

#### Check assumptions

```{r}
autoplot(plant.mod.no.na)
```

Again, don't worry if the assumptions aren't perfect. How do they compare to the first analysis? All of the plots look to be pretty identical to one another.

#### Interpret model

Use `anova()` first then `summary()`.

```{r}
anova(plant.mod.no.na)
summary(plant.mod.no.na)
```

Answer the following questions:

1.  How many degrees of freedom are there for Growth_rate_mm? Why?
  There is 1 degree of freedom for Growth_rate_mm because there are two groups of variables so you have to take the total number of groups subtracted by 1. So you would get 2-1=1.
2.  How many degrees of freedom are there for Residuals? Why?
  There are 34 degrees of freedom for the Residuals because, there are 36 total observations for this data set, and you have to take the total number of observations and subtract it by the number of variables, which would be 36-2=34.
3.  How are the degrees of freedom in this second model different than the first? What does the answer tell you about the role of NA values in the regression analysis?
  In the first model there was 60 observations in the data set, but we saw that 24 were taken out for "missingness" meaning for being NA values. We see that for the second model there were only 26 observations because I created a dataframe that took out all of the NAs. NA values cause there to be a "missing" number from the data observations which would cause it to be taken out of the data when running statistical analysis, but instead of making it more complicated its easier to just get ride of the NAs originially in a new data frame so you don't have to worry about any "missing" values in your data when running statistical analysis.
