---
title: "Homework-25"
author: "Margaret Dougher"
format: html
editor: visual
---

## Setting things up
```{r}
rm(list=ls()) #clears R's brain
library(here) #installing the necessary packages
library(tidyverse)
library(ggfortify)
```

## Importing data
```{r}
daphnia <- read.csv(here("Data", "Daphniagrowth (1).csv"))
glimpse(daphnia)
```

## Changing variables
```{r}
daphnia$parasite <- as.factor(daphnia$parasite)
glimpse(daphnia)
```

## Plotting and examining the data
```{r}
ggplot(daphnia, aes(x = parasite, y = growth.rate)) +
  geom_boxplot() +
  theme_bw()
```

## Manipulating the axis
```{r}
ggplot(daphnia, aes(x = parasite, y = growth.rate)) +
  geom_boxplot() +
  theme_bw() +
  coord_flip() #switches the x and y axis
```

## Constructing the ANOVA test
```{r}
model_grow <- lm(growth.rate ~ parasite, data = daphnia)
```

## Checking assumptions
```{r}
autoplot(model_grow, smooth.colour = NA)
```

## Making an inference from a one-way ANOVA
```{r}
anova(model_grow)
```

## Learning about treatment contrasts
```{r}
summary(model_grow)
```

A good tip i learned was the R likes to use the alphabet, so in the ANOVA framework we can assume that the word "Intercept" represents the first level of the treatment levels alphabetically labeled.

Treatment contrasts report differences between the reference level and the other levels.

# Getting the mean growth rates
```{r}
sumDat <- daphnia |>
  group_by(parasite) |>
  summarise(meanGR = mean(growth.rate))
```