---
title: "Homework-18"
author: "Margaret Dougher"
format: html
editor: visual
---

## Set things up

```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(dplyr)
```

## Importing data

```{r}
compensation <- read.csv(here("Data", "compensation.csv"))
```

## Graphing data

```{r}
ggplot(compensation, aes(x = Root, y = Fruit)) + 
  geom_point()
```

## Scatter plot tutorial

Plotting basics with ggplot

```{r}
rm(list=ls())
library(here)
library(dplyr)
library(ggplot2)
```

## Importing data

```{r}
compensation <- read.csv(here("Data", "compensation.csv"))
```

Looking at data

```{r}
glimpse(compensation)
```

Making my first ggplot picture

```{r}
ggplot(compensation, aes(x = Root, y = Fruit)) +
  geom_point() #creating the graph 

ggplot(compensation, aes(x = Root, y = Fruit)) +
  geom_point() +
  theme_bw() #getting ride of grey background on the graph

ggplot(compensation, aes(x = Root, y = Fruit)) +
  geom_point(size = 5) +
  theme_bw() #increasing the size of the points
```

Altering the x and y axis labels

```{r}
ggplot(compensation, aes(x = Root, y = Fruit)) +
  geom_point(size = 5) +
  xlab("Root Biomass") +
  ylab("Fruit Production") +
  theme_bw() #changing the x and y axis labels
```

Adjusting the colors of the points to match specific levels in a group

```{r}
ggplot(compensation, aes(x = Root, y = Fruit, colour = Grazing)) +
  geom_point(size = 5) +
  xlab("Root Biomass") +
  ylab("Fruit Production") +
  theme_bw()

ggplot(compensation, aes(x = Root, y = Fruit, shape = Grazing)) +
  geom_point(size = 5) +
  xlab("Root Biomass") +
  ylab("Fruit Production") +
  theme_bw() #changing the shapes of the points on the graph
```

Creating a box-and-whisker plot

```{r}
ggplot(compensation, aes(x = Grazing, y = Fruit)) +
  geom_boxplot() +
  xlab("Grazing treatment") +
  ylab("Fruit Production") +
  theme_bw()
```

Changing the size, colour, and transparency of the points

```{r}
ggplot(compensation, aes(x = Grazing, y = Fruit)) +
  geom_boxplot() +
  geom_point(size = 4, colour = 'lightgrey', alpha = 0.5) +
  xlab("Grazing treatment") +
  ylab("Fruit Production") +
  theme_bw()
```

Making a histogram of fruit production

```{r}
ggplot(compensation, aes(x = Fruit)) +
  geom_histogram()

# changing the binwidth

ggplot(compensation, aes(x = Fruit)) +
  geom_histogram(bins = 10)
ggplot(compensation, aes(x = Fruit)) +
  geom_histogram(binwidth = 15)
```

Learning how to produce a matrix of graphs automatically structered by a factor/categorical variable in your data

```{r}
ggplot(compensation, aes(x = Fruit)) +
  geom_histogram(binwidth = 15) +
  facet_wrap(~Grazing)
```
Saving graphs

```{r}
ggsave("ThatCoolHistogramOfFruit.png")
```


