---
title: "Homework-23"
author: "Margaret Dougher"
format: html
editor: visual
---

## Setting things up
```{r}
rm(list=ls()) #clear R's brain
library(here)
library(tidyverse) #installing necessary packages
```

## Importing data
```{r}
plants <- read.csv(here("Data", "plant.growth.rate.csv"))
glimpse(plants) #viewing data
```

## Plotting and examining the data
```{r}
ggplot(plants, aes(x = soil.moisture.content, y = plant.growth.rate)) +
  geom_point() +
  ylab("Plant Growth Rate (mm/week)") +
  theme_bw() #creating a scatterplot
```
## Making a linear regression
```{r}
model_pgr <- lm(plant.growth.rate ~ soil.moisture.content, data = plants)
```

## Making assumptions
```{r}
library(ggfortify)
autoplot(model_pgr, smooth.colour = NA)
```
We can see from the following plots that there is no pattern in either of the left-hand plots. The normal distribution assumption is met, and there are no points exerting high influence.

## Interpreting the data with the use of anova tables and summary tables
These tests will help us determine whether or not to reject the null hypothesis, and if our previous assumptions about the data are correct.

```{r}
anova(model_pgr)
summary(model_pgr)
```

## Using ggplot2 to add regression lines
```{r}
ggplot(plants, aes(x = soil.moisture.content, y = plant.growth.rate)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  ylab("Plant Growth Rate (mm/week)") +
  theme_bw()
```

