---
title: "Homework-21"
author: "Margaret Dougher"
format: html
editor: visual
---

## My first Chi-square contingency analysis

## Clear the decks
```{r}
rm(list=ls()) #clears R's brain
```
## Libraries I always use.
```{r}
library(dplyr)
library(ggplot2)
library(here)
```
## Importing the data
```{r}
lady <- read.csv(here("Data", "ladybirds.csv"))
glimpse(lady) #looking at the data
```
## Changing all character variables to factor variables to match the data in Chapter 5
```{r}
lady$Habitat <- as.factor(lady$Habitat)
lady$Site <- as.factor(lady$Site)
lady$colour <- as.factor(lady$colour)
```
## Changing all variables to start with uppercase letters
```{r}
colnames(lady)[3] = "Colour"
colnames(lady)[4] = "Number"
```
## Organizing the data for plotting and analysis
```{r}
totals <- lady |>
  group_by(Habitat, Colour) |>
  summarise(total.Number = sum(Number))
```
## Using a bar chart to summarise raw data
```{r}
ggplot(totals, aes(x = Habitat, y = total.Number, fill = Colour)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  scale_fill_manual(values = c(black = "black", red = "red"))
```
## Transforming data into the matrix needed to perform chi-square contingency test
```{r}
lady.mat <- xtabs(Number ~ Habitat + Colour, data = lady)
```
## Running the chi-square test
```{r}
lady.chi <- chisq.test(lady.mat) 
names(lady.chi)
```

## My first two sample t-test
```{r}
rm(list=ls()) #clearing R's brain
library(here)
library(dplyr)
library(ggplot2)
```
## Importing data
```{r}
ozone <- read.csv(here("Data", "ozone.csv"))
glimpse(ozone) #examining data
```
## Changing character variables to factor variables to match the data in Chapter 5
```{r}
ozone$Garden.location <- as.factor(ozone$Garden.location)
ozone$Garden.ID <- as.factor(ozone$Garden.ID)
```
## Creating a plot of data
```{r}
ggplot(ozone, aes(x = Ozone)) +
  geom_histogram(binwidth = 10) +
  facet_wrap(~ Garden.location, ncol = 1) + #breaking data into groups
  theme_bw()
```
## Using group_by and summarise functions to determine whether or not we should reject the null hypothesis
```{r}
totals <- ozone |>
  group_by(Garden.location, Garden.ID) |>
  summarise(total.Ozone = sum(Ozone))
```
## Performing a t-test
```{r}
t.test(Ozone ~ Garden.location, data = ozone)
```
## Comparing var-test function to t-test function
```{r}
var.test(Ozone ~ Garden.location, data = ozone)
```



